Writing Trash Bin

### 2024-02-26 paper.qmd
some word vomit of noise and signal, also some added citations

## Noise and Signal in Uncertainty Visualisation
Uncertainty is noise, not signal. While this comment may seem obvious, what may not be obvious is that the value abstraction experiments are treating uncertainty as *signal*. That is why the experimental results are identical to the existing information visualisation literature.

@@

## Bad Question stuff?
In the discussion of the HOPs, error bar, and violin plot study @Hullman2015 note:

> "In light of the common usage of error bars for presenting multiple independent distributions, it is noteworthy how poorly subjects using these representations performed on tasks asking them to assess how reliably a draw from one variable was larger than the other...Many subjects reported values less than 50%, which are not plausible given that the mean of B was larger than the mean of A."

>>>>

## Noise and Signal in Uncertainty Visualisation
Uncertainty is noise. This may be an obvious statement, but what is not obvious is that the existing literature sees noise as a signal. When we directly ask "What is the varinance of this variable" the variance *becomes* the signal. Noise and signal are not an inherrent property of variance, or meta-data or the aspects of an analysis that are usually secondary considerations, it is a property of the data that *promotes* a message we are trying to infer and the data that *supresses* it. The replication problem in uncertainty visualisation, where experiments are simply repeating established visualisation work, comes from the misunderstandings established in previous sections. Uncertainty is poorly defined, therefore authors do not understand it is not an inherrent property of the data and is actually related to a particular inference question, this incorrect view that uncertainty is always variance or probabilities leads to the class of "uncertainty visualisations" that depict different mathematical objects or visual features with little in the sense of consistent rules, this complicated class has caused visualisation authors to miss that they are simply repeating the established evaluation results from normal information visualisations. This situation is a bit of a mess. It also highlights a unique and facinating problem faced by uncertainty visualisation. If asking direct questions about uncertainty causes us to treat it as a signal, how do we evaluate uncertainty as *noise*?

Uncertainty evaluation studies are not all value extraction. There are cases where authors *know* the uncertainty information is relevant, but are not quite sure how to capture it. These studies most commonly evaluate a visualiation on its ability to convey information for a decision or how well it induces trust.

*decision and trust paragraph*

There are other studies that try to include uncertainty information without directly asking for it. This is an interesting approach, but these papers are often still looking for a "signal" which is specified in the form of a "ground truth" that is used to evaluate the plot. Unfortunately this usually comes in the form of slightly cryptic questions that create a large amount of noise on the interpretation side [@Hullman2016].
- @Hofmann2012 showed two distributions in 20 different visualisations (a lineup protocol) using a jittered sample, a density plot, a histogram, and a box plot. Participants were asked to report in which of the plots was "the blue group furthest to the right" and to provide reasoning from a multiple choice list and grade their certainty. The experiment set up is shown in @fig-right. The participants answers were then compared to a ground truth where the correct plot had a right shifted mean. By comparing the results to a ground truth statistic and marking participants as "wrong" or "right", the error from the participants that had an alternative interpretation to the concept of "furthest right" was conflated with the error from a poor design.

![This shows the user interface for the experiment performed by @Hofmann2012. The question of "furthest to the right" is open to interpretation. ](furthestright.png){#fig-right}

Additionally, the use of a ground truth highlights how participants were supposed to *use* the uncertainty information. In this informa

@@

### 2024-02-26 writingdumpster.qmd
some word vomit of noise and signal, also some added citations

# Main section outtakes
While mathematical information *should* be considered the starting point in considering if two visualisations are equivalent, it is not the only consideration. A large amount of work in perceptual tasks, attention, and psychology surrounding charts shows that all the information depicted is not paid attention to in equal measure. Elements such as colour can have sizable impacts on the way a graphic is percieved. Mathematics identifies the information that is *shown* while psychology identifies the difference in information *recieved*. A failure to understand this distinction leads to graphics that are different not only in the way information is depicted but in the information itself. This problem is pervasive in uncertainty visualisation literature.

>>>>

# Main section outtakes
Visualisation experiments are highly sensitive to the way questions are asked. For example, a study by @Hullman2015 that compared HOPs, error bar, and violin plots found that many participants reported that group A was more likely to be higher than group B, which was not not plausible given that the mean of B was larger than the mean of A. The authors attributed this obvious mistake to the visual representation, but we believe it was more likely caused by the distributions being shown in the order A then B while the question swapped the order by asking "What is P(B>A)?". This flip is minor, but must be done *after* the participants have worked out the probability. Since the primary task is done they forget to complete the secondary task, similar to forgetting your bank card in an ATM. I suspect this was the cause of the mistake as I, when attempting this question myself, made the exact same mistake. I also constalty lead my bank card in ATMs.

While mathematical information *should* be considered the starting point in considering if two visualisations are equivalent, it is not the only consideration. A large amount of work in perceptual tasks, attention, and psychology surrounding charts shows that all the information depicted is not paid attention to in equal measure. Elements such as colour can have sizable impacts on the way a graphic is percieved. Mathematics identifies the information that is *shown* while psychology identifies the difference in information *recieved*. A failure to understand this distinction leads to graphics that are different not only in the way information is depicted but in the information itself. This problem is pervasive in uncertainty visualisation literature.
